30 years of privacy research for the masses \& \newline 15 years of decentralization attempts

### 2014-03-13 paper.tex
Fix indents, extra \\ and other layour inconsistencies

\section{Introduction to Tor}
	\label{sec:tor}
	The implementation of The Onion Router was first described by the U.S. Navy Research Laboratory, as means to protect government communications from digital, as well as physical attacks, by hiding the location of the communicating party or parties \cite{goldschlag1996hiding}.

@@

		The routing in the network is done by Onion Routers, also called relay nodes. The relay nodes relays the data from the Onion Proxy to the web server across a circuit (circuits are described in 2D). Each Onion Router is connected to every other Onion Router with a TLS connection \cite{tlsprotocol}. Each circuit has three type of Onion Routers \cite{mccoy2008shining}:
		\begin{itemize}
			\item{The entrance Tor router:} this router is directly connected to an Onion Proxy and can observe the origin of a request through the Tor network. The entrance router sends the packet to the middle Tor router.
			\item{The middle Tor router:} this router is connected to the entrance router and the exit router.
			\item{The exit Tor router:} this router is connected to the web server. Note that the exit Tor router is the only router that can observe the final destination of the request.
		\end{itemize}

@@

	\subsection{Circuit creation}
		\label{ss:tor_circuit}
		As described in the previous section, data on the Tor networks travels over circuits. The data travels over the circuit in fixed-size cells that are 512 bytes long. Each cell has a header and a payload. The header consists of a number that identifies the circuit of the cell and a command that indicates what to do with the cell’s payload.

@@

		\subsection{Browser based attacks}
		Traffic analysis can be used to attack the anonymity of a user (a Tor client) browsing the web using Tor \cite{abbott2007browser}. By misusing the exit policy of Tor one can reduce the time required to perform the analysis.\\

		By running a HTTP exit node and an Tor router that eventually will act as an entry node in the network, an adversary can discover the identity of a user. The exit node injects an invisible iframe containing some JavaScript into the web pages that pass by, each sending a unique ID to a malicious web server. Every ten minutes the Tor client chooses a new circuit and eventually an unlucky Tor client picks and uses the malicious entry node that was placed in the network.\\

		By performing traffic analysis to compare the unique IDs of the web server and the circuits passing trough the entry node, a user can be identified via matches. A similar system can be set up only using the HTML meta refresh tag. To increase the odds of a user choosing the malicious exit node, one can run the exit node on unpopular ports. There are usually only a few exit nodes running on file sharing ports 4661 to 4666.\\

@@

			\subsection{Bootstrapping new nodes}
		If a Tor user decides to donate some of his bandwidth by running a bridge or a relay and thus creating a new node in the network, there has to be one starting point where this new node can discover neighbors in the network to connect with. In Tor a directory server can tell the new node what his neighbors are and how to find them \cite{dingledine2004tor}.\\

@@

		\subsection{Incentives in decentralized systems}
		As described in section \ref{ss:tor_disadvantages} people are using more bandwidth than that they donate. There are multiple proposals \cite{dingledine2010building, jansen13lira} to introduce incentives into Tor. If one would also have to build an incentive system into a decentralized system, they would have to find a way to manage the ratings of each client in the network, in such a way that they cannot be falsely modified. In other words the incentive data has to be accurate and reliable. Besides the integrity of the data, also the traffic it generates on the network should have minimal impact.\\

		Rahman \cite{rahman2009survey} proposes several options to build in incentives in a peer-to-peer network. The first thing that is described is the so called \emph{Warm-glow Model}. In this model determines the percentage of free-rides based on the probabilistic population distribution. If the percentage is above a certain threshold, the system will show signs of diminishing marginal returns.\\

		The second proposal is using monetary schemes. While it is not exemplified a lot, the main idea is to use a virtual currency as incentive. The problems with this approach are the scalability and the hidden costs of this service.\\

@@

	\section{Comparison existing decentralized networks}

		In the last section, we discussedexisting decentralized peer-to-peer networks. In this section we will make a comparison between these networks. For each network, we will explain the following features:
\begin{itemize}
\item Compatibility with Tor. If the network is compatible with Tor, the network or some features of the network could be used for the decentralization of Tor.
\item Public implementation. For each network we will explain whether a public implementation is available.
\item Used in practice. Is the network used in practice?
\item Attack resistance. We will explain the various weaknesses and possible attacks of each network.
\end{itemize}

	\section{Conclusion}
		In this technical report we introduced Tor and explained how Tor works and which components define a Tor network. Furthermore we looked at the main disadvantages and problems Tor currently has. We then made a bridge to decentralized systems. We first explained which problems one might face when decentralizing a system such as Tor. The next section shows which systems currently have a decentralized structure and/or mechanism. These systems are then compared to each other with their pros and cons.\\

		From this table we come to the conclusion that there is no complete decentralized system that provides the full functionality and security that Tor offers today. While there is a lot of research on decentralized systems, there is not a system yet that can compete with Tor. The systems that do exist such as Tarzan, Torsk and Gnutella are getting things right, yet they often lack functionality such as anonymity, good routing performance or are vulnerable for a variety of  attacks.\\

>>>>

\section{Introduction to Tor}
	\label{sec:tor}

	The implementation of The Onion Router was first described by the U.S. Navy Research Laboratory, as means to protect government communications from digital, as well as physical attacks, by hiding the location of the communicating party or parties \cite{goldschlag1996hiding}.

@@

		The routing in the network is done by Onion Routers, also called relay nodes. The relay nodes relays the data from the Onion Proxy to the web server across a circuit (circuits are described in 2D). Each Onion Router is connected to every other Onion Router with a TLS connection \cite{tlsprotocol}. Each circuit has three type of Onion Routers \cite{mccoy2008shining}:

		\begin{itemize}
			\item{The entrance Tor router:} this router is directly connected to an Onion Proxy and can observe the origin of a request through the Tor network. The entrance router sends the packet to the middle Tor router.
			\item{The middle Tor router:} this router is connected to the entrance router and the exit router.
			\item{The exit Tor router:} this router is connected to the web server. Note that the exit Tor router is the only router that can observe the final destination of the request.
		\end{itemize}

@@

	\subsection{Circuit creation}
		\label{ss:tor_circuit}

		As described in the previous section, data on the Tor networks travels over circuits. The data travels over the circuit in fixed-size cells that are 512 bytes long. Each cell has a header and a payload. The header consists of a number that identifies the circuit of the cell and a command that indicates what to do with the cell’s payload.

@@

		\subsection{Browser based attacks}
			Traffic analysis can be used to attack the anonymity of a user (a Tor client) browsing the web using Tor \cite{abbott2007browser}. By misusing the exit policy of Tor one can reduce the time required to perform the analysis.

			By running a HTTP exit node and an Tor router that eventually will act as an entry node in the network, an adversary can discover the identity of a user. The exit node injects an invisible iframe containing some JavaScript into the web pages that pass by, each sending a unique ID to a malicious web server. Every ten minutes the Tor client chooses a new circuit and eventually an unlucky Tor client picks and uses the malicious entry node that was placed in the network.

			By performing traffic analysis to compare the unique IDs of the web server and the circuits passing trough the entry node, a user can be identified via matches. A similar system can be set up only using the HTML meta refresh tag. To increase the odds of a user choosing the malicious exit node, one can run the exit node on unpopular ports. There are usually only a few exit nodes running on file sharing ports 4661 to 4666.

@@

		\subsection{Bootstrapping new nodes}
			If a Tor user decides to donate some of his bandwidth by running a bridge or a relay and thus creating a new node in the network, there has to be one starting point where this new node can discover neighbors in the network to connect with. In Tor a directory server can tell the new node what his neighbors are and how to find them \cite{dingledine2004tor}.

@@

		\subsection{Incentives in decentralized systems}
			As described in section \ref{ss:tor_disadvantages} people are using more bandwidth than that they donate. There are multiple proposals \cite{dingledine2010building, jansen13lira} to introduce incentives into Tor. If one would also have to build an incentive system into a decentralized system, they would have to find a way to manage the ratings of each client in the network, in such a way that they cannot be falsely modified. In other words the incentive data has to be accurate and reliable. Besides the integrity of the data, also the traffic it generates on the network should have minimal impact.

			Rahman \cite{rahman2009survey} proposes several options to build in incentives in a peer-to-peer network. The first thing that is described is the so called \emph{Warm-glow Model}. In this model determines the percentage of free-rides based on the probabilistic population distribution. If the percentage is above a certain threshold, the system will show signs of diminishing marginal returns.

			The second proposal is using monetary schemes. While it is not exemplified a lot, the main idea is to use a virtual currency as incentive. The problems with this approach are the scalability and the hidden costs of this service.

@@

	\section{Comparison existing decentralized networks}
		In the last section, we discussedexisting decentralized peer-to-peer networks. In this section we will make a comparison between these networks. For each network, we will explain the following features:
\begin{itemize}
\item Compatibility with Tor. If the network is compatible with Tor, the network or some features of the network could be used for the decentralization of Tor.
\item Public implementation. For each network we will explain whether a public implementation is available.
\item Used in practice. Is the network used in practice?
\item Attack resistance. We will explain the various weaknesses and possible attacks of each network.
\end{itemize}

	\section{Conclusion}
		In this technical report we introduced Tor and explained how Tor works and which components define a Tor network. Furthermore we looked at the main disadvantages and problems Tor currently has. We then made a bridge to decentralized systems. We first explained which problems one might face when decentralizing a system such as Tor. The next section shows which systems currently have a decentralized structure and/or mechanism. These systems are then compared to each other with their pros and cons.

		From this table we come to the conclusion that there is no complete decentralized system that provides the full functionality and security that Tor offers today. While there is a lot of research on decentralized systems, there is not a system yet that can compete with Tor. The systems that do exist such as Tarzan, Torsk and Gnutella are getting things right, yet they often lack functionality such as anonymity, good routing performance or are vulnerable for a variety of  attacks.
